---
title: "Data Guide Creation and Level-0 Data Compilation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_guide_and_LO}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

The `invitroTKstats` R package is designed to work as a data processing pipeline for *in vitro* assays collecting various physiological parameters used in high-throughput toxicokinetic (HTTK) modeling.  Mass spectrometry (MS) data used to estimate the various physiological parameters.

Here, we discuss the first steps of the data processing pipeline, which include the data guide creation and level-0 (raw data) compilation.

## Suggested packages for use with this vignette

```{r setup}
# Primary Packages #
library(invitroTKstats)
# Data Formatting Packages #
library(dplyr)
library(magrittr)
library(stringr)
```

# Data Catalog Creation

The data catalog creation step can be thought of as a log of all the relevant raw data files for a given physiological parameter that we wish to pipeline.  It is a crucial part of the pipeline in that it allows us to identify where all of the raw datafiles are coming from and where we are pulling raw (Level-0) data from within those datasets.  This provides key meta-data information that can be used for evaluating data provenance.

Though this `data.frame` may be created manually through an Excel file, the `invitroTKstats` package includes a function called `create_data_catalog` that will automatically generate and document the creation of the data catalog.  This allows for maximum reproducibility, transparency, and efficiency in compiling the necessary data guide.  Thus, it is a best practice to utilize this function for creating the data file.

Below is an example of how to use this function

```{r}
# create a data catalog listing where Fup data is held
DC <- create_catalog(
  file = "<file_path>/my_raw_MS_data_Clint_MMDDYYYY.xlsx",
  sheet = c(rep(1,5),rep(2,5),rep(3,5)),
  skip.rows = rep(c(1,10,19,28,37),3),
  num.rows = 8,date = "MMDDYYYY"
  )
```


# Session Information

```{r}
sessionInfo()
```